---
title: '**Reproducing a model**'
author: "Eleanor Burch (1), Ellen Brooks-Pollock (1), and Hannah Christensen (1)

(1) Population Health Sciences, Bristol Medical School, University of Bristol, Bristol, United Kingdom.
"
date: "26/09/2022"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE,comment = NA, out.width='600px', dpi=200)
```


Here we reproduced and extended a mathematical model of SARS-CoV-2 vaccine prioritisation strategies developed by Hill and Keeling.(1)  We have introduced life-years saved as an alternative outcome to deaths averted. We also included up-to-date estimates for the risk of mortality due to COVID-19 and vaccine uptake levels for different groups of the population of England.


#### **Priority groups**

The JCVI priority groups for vaccination are as follows(2):  
1. Care homes  
2. ≥80 years & Health Care Workers (HCW)  
3. 75–79 years  
4. 70–74 years & Clinically Extremely Vulnerable (CEV)  
5. 65–69 years  
6. Underlying Health Conditions (UHC)  
7. 60–64 years  
8. 55–59 years  
9. 50–54 years  
10. 0–49 years  


The estimates for the sizes of each priority group in the population (in millions, P) are(3):

```{r}

P <- c(0.7,6.0,1.9,3.7,2.4,6.2,1.5,2.0,2.3,29.3)      

```

#### **Plotting deaths averted against number of available doses**  

v1 represents the proportion of each group that will receive one dose of vaccine, with the initial values set to zero.

```{r}  

v1 <- c(0,0,0,0,0,0,0,0,0,0)         

```            

v2 represents the proportion of each group that will receive two doses of vaccine, with the initial values set to zero.

```{r}  

v2 <- c(0,0,0,0,0,0,0,0,0,0)       

```            

The number of doses that have been allocated is initially set to zero.

```{r}

dosesallocated <- 0   

```

The values for vaccine efficacy were taken directly from Hill and Keeling, and were based on estimates from Public Health England of efficacy of both the Pfizer and AstraZeneca vaccines against COVID-19 mortality, in relation to the Alpha variant.(4) Here, we took the upper and lower estimates of the vaccine efficacy (70 – 85% for one dose and 75 – 99% for two doses) and a range of values in this interval were considered.

The vaccine efficacy after the first dose is varied from 70% to 85% in steps of 5%:

```{r}

VE1 <- seq(0.7,0.85,0.05)

```

The vaccine efficacy after the second dose is varied from 75% to 99% in steps of 8%:

```{r}

VE2 <- seq(0.75,0.99,0.08)  

```

The number of available vaccine doses is initially varied from 0 to 4 million in steps of 1 million:

```{r}

dosesavailable <- seq(0,4,1) 

```

SARS-CoV-2 vaccine uptake levels for the age-groups, care home residents and staff, healthcare workers, and people who are clinically extremely vulnerable were obtained from the National Health Service (NHS) England website, using the ‘weekly announced vaccinations’ data from 3rd March 2022.(5) Vaccine uptake levels for people with underlying health conditions were assumed to be an average of the uptake levels for those aged 60-64 years and those aged 65-69 years.

Uptake levels of the first dose for each priority groups are as follows:

```{r}

U1 <- c(0.96,0.93,1.0,0.96,0.97,0.98,0.99,0.98,0.95,0.84)     

```

Uptake levels of the second dose for each priority groups are as follows:

```{r}

U2 <- c(0.95,0.91,1.0,0.95,0.96,0.97,0.98,0.96,0.93,0.75)       

```

The risk of mortality for the age-groups was estimated using the number of deaths registered where COVID-19 was mentioned on the death certificate in each group before the vaccination programme began, using the 2020 Office for National Statistics (ONS) dataset showing deaths registered weekly in England and Wales during the period 13th March – 11th December 2020.(6) These death counts, for nine months of the year, were then multiplied by 4/3 ̇ to estimate the number of deaths that would have occurred in one year without vaccination. The number of deaths was also scaled using the proportion of total deaths in the period that occurred in England compared to in both England and Wales, in order to give estimates for the number of deaths that occurred just in England. The mortality risk was calculated by dividing the number of deaths in each age-group by the population size of the group. The mortality risk for the care homes group was estimated through the same method, using the 2020 ONS dataset showing deaths of care home residents notified to the Care Quality Commission (CQC), during the period 10th April – 18th December 2020.(7) The mortality risk for care home staff has been included in this estimate and is assumed to be the same as that for care home residents. The mortality risks for healthcare workers and people who are clinically extremely vulnerable were assumed to be the same as for the age range they are grouped with in the priority group structure (≥80 years and 70-74 years, respectively). The mortality risk for those with underlying health conditions was assumed to be an average for those aged 60-64 years and those aged 65-69 years.

Estimates for the risk of mortality due to COVID-19 for each priority group (R) are as follows:

```{r}  

R <- c(0.038738571,0.009145663,0.006533286,0.002312164,0.002134561,0.00228254,0.002430518,0.001236421,0.000629129,0.0000586534)

```

Using these vaccine uptake levels and risk of mortality, the number of deaths averted are plotted against the total number of available doses, following the one-dose only and two-dose only allocation strategies. *Figure 1* shows the number of deaths averted with up to four million available doses. At around 1.5 million doses, with a second-dose vaccine efficacy of 99% and a first-dose efficacy of 70%, the two-dose strategy briefly takes over as the optimal strategy. It then switches back at around 3 million doses to the one-dose strategy.

```{r, echo=FALSE, fig.cap = "Figure 1. Number of deaths averted following the one-dose only and two-dose only SARS-CoV-2 vaccination strategies, versus the number of available doses, with up to four million doses available. The blue lines correspond to the one-dose strategy, and the red lines correspond to the two-dose strategy, with a range of vaccine efficacies considered for each strategy."}


# One-dose vaccine allocation strategy

dosesavailable <- seq(0,4,1) 
deathsaverted <- dosesavailable
blue_colour_range <- colorRampPalette(c("lightblue","blue"))
blues <- blue_colour_range(4)
red_colour_range <- colorRampPalette(c("coral","darkred"))
reds <- red_colour_range(4)
plotnum <- 1
allocation.matrix <- matrix(0,nrow=26,ncol=10)

plot(1, type="n", xlab="Doses available (millions)", ylab="Deaths averted",xlim=c(0, 4), ylim = c(0, 50000))

legend("bottomright", 
       legend = c("1-dose, VE1 = 0.7","1-dose, VE1 = 0.75","1-dose, VE1 = 0.80","1-dose, VE1 = 0.85", "2-dose, VE2 = 0.75","2-dose, VE2 = 0.83","2-dose, VE2 = 0.91","2-dose, VE2 = 0.99"), 
       col = c(blues,reds),  
       pch = 16, 
       bty = "n", 
       pt.cex = 1, 
       cex = 0.8, 
       text.col = "black", 
       inset = c(0.01, 0.001))

# The following loop allocates the doses in order from the highest priority group down to the lowest priority group, given a fixed number of available doses (in millions), and according to the vaccine uptake levels above. The doses are allocated following the one-dose strategy

# Equation 1 calculates the annual number of deaths averted following the one-dose allocation strategy 


for(k in 1:length(VE1)){
  for(j in 1:length(dosesavailable)){ 
    v1 <- c(0,0,0,0,0,0,0,0,0,0)
    dosesallocated <- 0
    for (i in 1:length(v1)){
      if (dosesavailable[j]>(dosesallocated+U1[i]*P[i])){
        v1[i] <- U1[i]
        dosesallocated <- dosesallocated + U1[i]*P[i]
      }
      else if ((dosesavailable[j]<(dosesallocated+U1[i]*P[i])) && (dosesavailable[j]>dosesallocated)){
        v1[i] <- (dosesavailable[j] - dosesallocated)/P[i]
        dosesallocated <- dosesallocated + v1[i]*P[i]
      }
      else {
        v1[i] <- 0
      }
    }
    deathsaverted[j] = sum(v1*VE1[k]*P*R*10^6)              #Equation 1
    allocation.matrix[j,] <- v1
  }
  lines(dosesavailable,deathsaverted, type = "l",col=blues[plotnum])
  plotnum <- plotnum + 1
}


# Two-dose vaccine allocation strategy

dosesallocated <- 0
deathsaverted <- dosesavailable
plotnum <- 1

# The following loop allocates the doses in order from the highest priority group down to the lowest priority group, given a fixed number of available doses (in millions), and according to the vaccine uptake levels above. The doses are allocated following the one-dose strategy

# Equation 2 calculates the annual number of deaths averted following the two-dose allocation strategy 

for(k in 1:length(VE2)){
  for(j in 1:length(dosesavailable)){
    v2 <- c(0,0,0,0,0,0,0,0,0,0)
    dosesallocated <- 0
    for (i in 1:length(v2)){
      if ((dosesavailable[j]/2)>(dosesallocated+U2[i]*P[i])){
        v2[i] <- U2[i]
        dosesallocated <- dosesallocated + U2[i]*P[i]
      }
      else if (((dosesavailable[j]/2)<(dosesallocated+U2[i]*P[i])) && ((dosesavailable[j]/2)>dosesallocated)){
        v2[i] <- ((dosesavailable[j]/2) - dosesallocated)/P[i]
        dosesallocated <- dosesallocated + v2[i]*P[i]
      }
      else {
        v2[i] <- 0
      }
    }
    deathsaverted[j] = sum(v2*VE2[k]*P*R*10^6)             #Equation 2
    allocation.matrix[j+length(dosesavailable),] <- v2
  }
  lines(dosesavailable,deathsaverted, type = "l",col=reds[plotnum])
  plotnum <- plotnum + 1
}



```

*Figure 2* shows the number of deaths averted for a higher number of available doses - up to 60 million.

```{r, echo=FALSE, fig.cap = "Figure 2. Number of deaths averted following the one-dose and two-dose SARS-CoV-2 vaccination strategies, versus the number of available doses, with up to 60 million doses. The blue lines correspond to the one-dose strategy, and the red lines correspond to the two-dose strategy, with a range of vaccine efficacies considered for each strategy."}


# One-dose vaccine allocation strategy

dosesavailable <- seq(0,60,5) 
v1 <- c(0,0,0,0,0,0,0,0,0,0)
dosesallocated <- 0
VE1 <- seq(0.7,0.85,0.05) 
deathsaverted <- dosesavailable
blue_colour_range <- colorRampPalette(c("lightblue","blue"))
blues <- blue_colour_range(4)
red_colour_range <- colorRampPalette(c("coral","darkred"))
reds <- red_colour_range(4)
plotnum <- 1
allocation.matrix <- matrix(0,nrow=26,ncol=10)

plot(1, type="n", xlab="Doses available (millions)", ylab="Deaths averted",xlim=c(0, 60), ylim = c(0, 130000))

legend("bottomright", 
       legend = c("1-dose, VE1 = 0.7","1-dose, VE1 = 0.75","1-dose, VE1 = 0.80","1-dose, VE1 = 0.85", "2-dose, VE2 = 0.75","2-dose, VE2 = 0.83","2-dose, VE2 = 0.91","2-dose, VE2 = 0.99"), 
       col = c(blues,reds),  
       pch = 16, 
       bty = "n", 
       pt.cex = 1, 
       cex = 0.8, 
       text.col = "black", 
       inset = c(0.01, 0.001))


for(k in 1:length(VE1)){
  for(j in 1:length(dosesavailable)){ 
    v1 <- c(0,0,0,0,0,0,0,0,0,0)
    dosesallocated <- 0
    for (i in 1:length(v1)){
      if (dosesavailable[j]>(dosesallocated+U1[i]*P[i])){
        v1[i] <- U1[i]
        dosesallocated <- dosesallocated + U1[i]*P[i]
      }
      else if ((dosesavailable[j]<(dosesallocated+U1[i]*P[i])) && (dosesavailable[j]>dosesallocated)){
        v1[i] <- (dosesavailable[j] - dosesallocated)/P[i]
        dosesallocated <- dosesallocated + v1[i]*P[i]
      }
      else {
        v1[i] <- 0
      }
    }
    deathsaverted[j] = sum(v1*VE1[k]*P*R*10^6)
    allocation.matrix[j,] <- v1
  }
  lines(dosesavailable,deathsaverted, type = "l",col=blues[plotnum])
  plotnum <- plotnum + 1
}


# Two-dose vaccine allocation strategy

dosesavailable <- seq(0,60,5) 
v2 <- c(0,0,0,0,0,0,0,0,0,0)
dosesallocated <- 0
VE2 <- seq(0.75,0.99,0.08)
deathsaverted <- dosesavailable
plotnum <- 1


for(k in 1:length(VE2)){
  for(j in 1:length(dosesavailable)){
    v2 <- c(0,0,0,0,0,0,0,0,0,0)
    dosesallocated <- 0
    for (i in 1:length(v2)){
      if ((dosesavailable[j]/2)>(dosesallocated+U2[i]*P[i])){
        v2[i] <- U2[i]
        dosesallocated <- dosesallocated + U2[i]*P[i]
      }
      else if (((dosesavailable[j]/2)<(dosesallocated+U2[i]*P[i])) && ((dosesavailable[j]/2)>dosesallocated)){
        v2[i] <- ((dosesavailable[j]/2) - dosesallocated)/P[i]
        dosesallocated <- dosesallocated + v2[i]*P[i]
      }
      else {
        v2[i] <- 0
      }
    }
    deathsaverted[j] = sum(v2*VE2[k]*P*R*10^6)
    allocation.matrix[j+length(dosesavailable),] <- v2
  }
  lines(dosesavailable,deathsaverted, type = "l",col=reds[plotnum])
  plotnum <- plotnum + 1
}


```

*Figures 3(a) - (m)* below show the proportions of each priority group that have been vaccinated following the one-dose only and two-dose only allocation strategies, with a total number of available doses ranging from 0 to 60 million.

```{r, echo=FALSE, fig.cap = "Figures 3(a) - (m). Proportions of each priority group that have been vaccinated following the one-dose only and two-dose only allocation strategies, with a total number of available doses ranging from 0 to 60 million. The blue bars correspond to the one-dose only strategy, and the red bars correspond to the two-dose only strategy."}

library(ggplot2)
library(patchwork)
library(gtable)
library(grid)
library(gridExtra)
library(cowplot)


prioritygroup <- c("1","2","3","4","5","6","7","8","9","10","1","2","3","4","5","6","7","8","9","10")
Dose_strategy <- c("1-dose", "1-dose", "1-dose", "1-dose", "1-dose", "1-dose", "1-dose", "1-dose", "1-dose", "1-dose", "2-dose", "2-dose", "2-dose", "2-dose", "2-dose", "2-dose", "2-dose", "2-dose", "2-dose", "2-dose")
plots <- list()

titles = array(c("a","b","c","d","e","f","g","h","i","j","k","l","m"))

for (i in 1:length(dosesavailable)){
  data <- data.frame(
    prioritygroup=prioritygroup,  
    proportionvaccinated=c(allocation.matrix[i,],allocation.matrix[i+length(dosesavailable),])
  )
  
plots[[i]] <- ggplot(data, aes(x=prioritygroup, y=proportionvaccinated, fill=Dose_strategy)) + 
  ggtitle(titles[[i]]) + 
  geom_col(position = "dodge") + 
  scale_x_discrete (limits = c("1","2","3","4","5","6","7","8","9","10"), name = "Priority group") +
  scale_y_continuous (name = "Proportion vaccinated",limits = c(0,1.0),breaks=c(0, 0.2, 0.4, 0.6, 0.8, 1.0)) +
  scale_fill_manual(values = c("cadetblue", "coral") ) +
  theme(plot.title=element_text(size=7),
        axis.text.x = element_text(size=5,hjust = 1, vjust = 1),
        axis.text.y = element_text(size=5), 
        axis.title = element_text(size = 5),
        plot.tag = element_text(size=5),
        plot.tag.position = "bottom",
        plot.margin = unit(c(0, 0.2, 0, 0.2), "cm"))
legend <- get_legend(plots[[i]])
plots[[i]] <- plots[[i]] + theme(legend.position="none")
}


plots[[14]] <- legend

wrap_plots(plots,ncol = 4)

```

#### **Estimating life-years saved as an alternative measure of disease impact**

The expected life-years remaining for the age-groups were calculated by taking the mean expected number of years remaining across the ages in the age range, using the ONS National Life Tables dataset consisting of period life expectancies for the UK population, based on data from the years 2018 - 2020.(8) The life-years remaining for care home residents and staff were assumed to be the same as for the over-80s. The life-years remaining for healthcare workers and people who are clinically extremely vulnerable were, as with the risk of mortality, assumed to be the same as for the age-groups they are grouped with in the priority group structure. The life-years remaining for those with underlying health conditions were assumed to be an average of the estimates for people aged 60-64 years and for people aged 65-69 years.

The expected number of years of life remaining for each priority group are below:

```{r}  

L <- c(4.74,4.74,10.86,14.34,18.16,20.19,22.22,26.49,30.93,57.16) 

```


*Figure 4(a)* shows the number of life-years saved with lower numbers of available doses - up to four million. Life-years saved are plotted against the total number of available doses, following the one-dose only and two-dose only allocation strategies. At around 1.5 million doses,  with a second-dose vaccine efficacy of 99% and a first-dose efficacy of 70%, the two-dose strategy briefly takes over as the optimal strategy. It then switches back at around 3 million doses to the one-dose strategy.

*Figure 4(b)* shows the number of life-years saved following each vaccination strategy with up to 60 million available doses.

```{r, echo=FALSE, fig.cap = "Figures 4(a) - (b). Number of life-years saved following the one-dose only and two-dose only SARS-CoV-2 vaccination strategies, versus the number of available doses. The blue lines correspond to the one-dose strategy, and the red lines correspond to the two-dose strategy, with a range of vaccine efficacies considered for each strategy."}

v1 <- c(0,0,0,0,0,0,0,0,0,0)
dosesallocated <- 0
VE1 <- seq(0.7,0.85,0.05)         
dosesavailable <- seq(0,4,1) 
lifeyearssaved <- dosesavailable
blue_colour_range <- colorRampPalette(c("lightblue","blue"))
blues <- blue_colour_range(4)
red_colour_range <- colorRampPalette(c("coral","darkred"))
reds <- red_colour_range(4)
plotnum <- 1
allocation.matrix <- matrix(0,nrow=26,ncol=10)

par(mfrow = c(1, 2))

plot(1, type="n", xlab="Doses available (millions)", ylab="Life-years saved", xlim=c(0, 4), ylim = c(0, 250000), main="a", font.main=1)


for(k in 1:length(VE1)){
  for(j in 1:length(dosesavailable)){
    v1 <- c(0,0,0,0,0,0,0,0,0,0)
    dosesallocated <- 0
    for (i in 1:length(v1)){
      if (dosesavailable[j]>(dosesallocated+U1[i]*P[i])){
        v1[i] <- U1[i]
        dosesallocated <- dosesallocated + U1[i]*P[i]
      }
      else if ((dosesavailable[j]<(dosesallocated+U1[i]*P[i])) && (dosesavailable[j]>dosesallocated)){
        v1[i] <- (dosesavailable[j] - dosesallocated)/P[i]
        dosesallocated <- dosesallocated + v1[i]*P[i]
      }
      else {
        v1[i] <- 0
      }
    }
    lifeyearssaved[j] = sum(v1*VE1[k]*P*R*L*10^6)                 # Equation 3
    allocation.matrix[j,] <- v1
  }
  lines(dosesavailable,lifeyearssaved, type = "l",col=blues[plotnum])
  plotnum <- plotnum + 1
}


# Two-dose vaccine allocation strategy

# Equation 4 calculates the annual number of deaths averted following the two-dose allocation strategy

v2 <- c(0,0,0,0,0,0,0,0,0,0)
dosesallocated <- 0
VE2 <- seq(0.75,0.99,0.08)          
lifeyearssaved <- dosesavailable
plotnum <- 1


for(k in 1:length(VE2)){
  for(j in 1:length(dosesavailable)){
    v2 <- c(0,0,0,0,0,0,0,0,0,0)
    dosesallocated <- 0
    for (i in 1:length(v2)){
      if ((dosesavailable[j]/2)>(dosesallocated+U2[i]*P[i])){
        v2[i] <- U2[i]
        dosesallocated <- dosesallocated + U2[i]*P[i]
      }
      else if (((dosesavailable[j]/2)<(dosesallocated+U2[i]*P[i])) && ((dosesavailable[j]/2)>dosesallocated)){
        v2[i] <- ((dosesavailable[j]/2) - dosesallocated)/P[i]
        dosesallocated <- dosesallocated + v2[i]*P[i]
      }
      else {
        v2[i] <- 0
      }
    }
    lifeyearssaved[j] = sum(v2*VE2[k]*P*R*L*10^6)                 # Equation 4
    allocation.matrix[j+length(dosesavailable),] <- v2
  }
  lines(dosesavailable,lifeyearssaved, type = "l",col=reds[plotnum])
  plotnum <- plotnum + 1
}


# One-dose vaccine allocation strategy

v1 <- c(0,0,0,0,0,0,0,0,0,0)
dosesallocated <- 0
VE1 <- seq(0.7,0.85,0.05)         
dosesavailable <- seq(0,60,5) 
lifeyearssaved <- dosesavailable
blue_colour_range <- colorRampPalette(c("lightblue","blue"))
blues <- blue_colour_range(4)
red_colour_range <- colorRampPalette(c("coral","darkred"))
reds <- red_colour_range(4)
plotnum <- 1
allocation.matrix <- matrix(0,nrow=26,ncol=10)


plot(1, type="n", xlab="Doses available (millions)", ylab="Life-years saved",xlim=c(0, 60), ylim = c(0, 1200000),main="b", font.main=1)

legend("bottomright", 
       legend = c("1-dose, VE1 = 0.7","1-dose, VE1 = 0.75","1-dose, VE1 = 0.80","1-dose, VE1 = 0.85", "2-dose, VE2 = 0.75","2-dose, VE2 = 0.83","2-dose, VE2 = 0.91","2-dose, VE2 = 0.99"), 
       col = c(blues,reds),  
       pch = 16, 
       bty = "n", 
       pt.cex = 0.8, 
       cex = 0.7, 
       text.col = "black", 
       inset = c(0.01, 0.001))

for(k in 1:length(VE1)){
  for(j in 1:length(dosesavailable)){
    v1 <- c(0,0,0,0,0,0,0,0,0,0)
    dosesallocated <- 0
    for (i in 1:length(v1)){
      if (dosesavailable[j]>(dosesallocated+U1[i]*P[i])){
        v1[i] <- U1[i]
        dosesallocated <- dosesallocated + U1[i]*P[i]
      }
      else if ((dosesavailable[j]<(dosesallocated+U1[i]*P[i])) && (dosesavailable[j]>dosesallocated)){
        v1[i] <- (dosesavailable[j] - dosesallocated)/P[i]
        dosesallocated <- dosesallocated + v1[i]*P[i]
      }
      else {
        v1[i] <- 0
      }
    }
    lifeyearssaved[j] = sum(v1*VE1[k]*P*R*L*10^6)           #Equation 3
    allocation.matrix[j,] <- v1
  }
  lines(dosesavailable,lifeyearssaved, type = "l",col=blues[plotnum])
  plotnum <- plotnum + 1
}


# Two-dose vaccine allocation strategy

v2 <- c(0,0,0,0,0,0,0,0,0,0)
dosesallocated <- 0
VE2 <- seq(0.75,0.99,0.08)          
lifeyearssaved <- dosesavailable
plotnum <- 1


for(k in 1:length(VE2)){
  for(j in 1:length(dosesavailable)){
    v2 <- c(0,0,0,0,0,0,0,0,0,0)
    dosesallocated <- 0
    for (i in 1:length(v2)){
      if ((dosesavailable[j]/2)>(dosesallocated+U2[i]*P[i])){
        v2[i] <- U2[i]
        dosesallocated <- dosesallocated + U2[i]*P[i]
      }
      else if (((dosesavailable[j]/2)<(dosesallocated+U2[i]*P[i])) && ((dosesavailable[j]/2)>dosesallocated)){
        v2[i] <- ((dosesavailable[j]/2) - dosesallocated)/P[i]
        dosesallocated <- dosesallocated + v2[i]*P[i]
      }
      else {
        v2[i] <- 0
      }
    }
    lifeyearssaved[j] = sum(v2*VE2[k]*P*R*L*10^6)             #Equation 4
    allocation.matrix[j+length(dosesavailable),] <- v2
  }
  lines(dosesavailable,lifeyearssaved, type = "l",col=reds[plotnum])
  plotnum <- plotnum + 1
}

```

### **Reference list**

1. Hill EM, Keeling MJ. Comparison between one and two dose SARS-CoV-2 vaccine prioritization for a fixed number of vaccine doses. J R Soc Interface [Internet]. 2021 Sep [cited 2022 Sep 25];18(182). Available from: https://royalsocietypublishing.org/doi/10.1098/rsif.2021.0214#d1e1364

2. Joint Committee on Vaccination and Immunisation: advice on priority groups for COVID-19 vaccination, 30 December 2020 - GOV.UK [Internet]. [cited 2022 Apr 29]. Available from: https://www.gov.uk/government/publications/priority-groups-for-coronavirus-covid-19-vaccination-advice-from-the-jcvi-30-december-2020/joint-committee-on-vaccination-and-immunisation-advice-on-priority-groups-for-covid-19-vaccination-30-december-2020

3. Department of Health and Social Care. UK COVID-19 vaccines delivery plan [Internet]. 2021 [cited 2022 Mar 17]. Available from: https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/951928/uk-covid-19-vaccines-delivery-plan-final.pdf

4. Public Health England. COVID-19 vaccine surveillance report: 1 July 2021 (week 26) [Internet]. 2021 [cited 2022 Mar 17]. Available from: https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/998411/Vaccine_surveillance_report_-_week_26.pdf

5. NHS England. COVID-19 Vaccinations: Weekly announced vaccinations data from 3rd March 2022 [Internet]. [cited 2022 Mar 17]. Available from: https://www.england.nhs.uk/statistics/statistical-work-areas/covid-19-vaccinations/

6. Deaths registered weekly in England and Wales, provisional - Office for National Statistics [Internet]. [cited 2022 Mar 17]. Available from: https://www.ons.gov.uk/peoplepopulationandcommunity/birthsdeathsandmarriages/deaths/datasets/weeklyprovisionalfiguresondeathsregisteredinenglandandwales

7. Number of deaths in care homes notified to the Care Quality Commission, England - Office for National Statistics [Internet]. [cited 2022 Mar 17]. Available from: https://www.ons.gov.uk/peoplepopulationandcommunity/birthsdeathsandmarriages/deaths/datasets/numberofdeathsincarehomesnotifiedtothecarequalitycommissionengland

8. National life tables: UK - Office for National Statistics [Internet]. [cited 2022 Mar 17]. Available from: https://www.ons.gov.uk/peoplepopulationandcommunity/birthsdeathsandmarriages/lifeexpectancies/datasets/nationallifetablesunitedkingdomreferencetables

